---
title: "Unmet Needs in the IBD Population"
author: "Gemma Postill"
date: "2022-07-29"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(tableone)
library(survey)
library(gtsummary)
library(pscl)
library(scales)
# library(MASS) # Note this package unmasks select from dplyr and wreaks havoc on the code
library(questionr)
library(pivottabler)
library(car)
library(mctest)
library(lme4)
library(mlmhelpr)

```

# Data Preparation

> Loading the data 

```{r setup1, include=FALSE}
#This line would need to be changed 


### IMPORT THE DATASET 
cchs2014 <- read_csv("CSV NAME")



```


### CLEANING: AGE & SEX VARIABLES 

```{r data cleaning: age and sex}
cchs20141 <- cchs2014 %>% 
  mutate(sex = factor(DHH_SEX,
                      levels = c(1, 2),
                      labels = c("Male", "Female")))

cchs20141 <- cchs20141 %>% 
  #This variable is grouping age in categories of 10-years 
  mutate(age_group = case_when(
    DHHGAGE == 1 | DHHGAGE == 2 | DHHGAGE == 3 ~ "< 20", #1 = 12-14, 2 = 15-17, 3 = 18-19
    DHHGAGE == 4 | DHHGAGE == 5 ~ "18 to 29", 
    DHHGAGE == 6 | DHHGAGE == 7 ~ "30 to 39",
    DHHGAGE == 8 | DHHGAGE == 9 ~ "40 to 49",
    DHHGAGE == 10 | DHHGAGE == 11 ~ "50 to 59",
    DHHGAGE == 12 | DHHGAGE == 13 ~ "60 to 69",
    DHHGAGE == 14 | DHHGAGE == 15 ~ "70 to 79",
    DHHGAGE == 16 ~ "80 or older"
  )) 

cchs20141 <- cchs20141 %>% 
  #This variable is grouping age in categories of 20-years 
  mutate(age_group_20s_OG = case_when(
    DHHGAGE == 1 | DHHGAGE == 2 | DHHGAGE == 3 ~ "< 20", #1 = 12-14, 2 = 15-17, 3 = 18-19
    DHHGAGE == 4 | DHHGAGE == 5 | DHHGAGE == 6 | DHHGAGE == 7 ~ "20 to 39",
    DHHGAGE == 8 | DHHGAGE == 9 | DHHGAGE == 10 | DHHGAGE == 11 ~ "40 to 59",
    DHHGAGE == 12 | DHHGAGE == 13 | DHHGAGE == 14 | DHHGAGE == 15 ~ "60 to 79",
    DHHGAGE == 16 ~ "80 or older"
  )) 

#Then we need to EXCLUDE the category of those with age_group = "< 20" & age_group_20s = "< 20" 
cchs20141<-subset(cchs20141, age_group!="< 20" & age_group_20s_OG!="< 20")

#Displays the results
cchs20141 %>% count(sex)
cchs20141 %>% count(age_group)
cchs20141 %>% count(age_group_20s_OG)

```

### CLEANING: IBD VARIABLE

```{r data cleaning: IBD1}
cchs20141 <- cchs20141 %>% 
  mutate(bowel_disease = case_when(
    CCC_171 == 1 & CCC_17A == 1 ~ "Crohn's disease",
    CCC_171 == 1 & CCC_17A == 2 ~ "Ulcerative colitis",
    CCC_171 == 1 & CCC_17A == 3 ~ "Irritable bowel syndrome",
    CCC_171 == 1 & CCC_17A == 4 ~ "Bowel incontinence",
    CCC_171 == 1 & CCC_17A == 5 ~ "Other bowel disease",
    CCC_171 == 2 ~ "No bowel disorder"
  )) %>% 
  mutate(ibd_yes = case_when(
    bowel_disease == "Crohn's disease" | bowel_disease == "Ulcerative colitis" ~ 1,
    TRUE ~ 0
  )) %>% 
  mutate(ibd = case_when(
    ibd_yes == 1 ~ "IBD Population", 
    ibd_yes == 0 ~ "General Population", 
  )) %>% 
  mutate(ibd_type = case_when(
    bowel_disease == "Crohn's disease" | bowel_disease == "Ulcerative colitis" ~ bowel_disease,
  ))

cchs20141 %>% count(bowel_disease)
cchs20141 %>% count(ibd)
cchs20141 %>% count(ibd_type)


#Then we need to EXCLUDE the category of those with:
  #bowel_disease != "Crohn's disease"
  #bowel_disease != "Ulcerative colitis"

#cchs20141<-subset(cchs20141, bowel_disease!="Irritable bowel syndrome" & bowel_disease!="Bowel incontinence"
#                  & bowel_disease!="Other bowel disease" )

cchs20141<-subset(cchs20141, bowel_disease!="Irritable bowel syndrome") 
cchs20141<-subset(cchs20141, bowel_disease!="Bowel incontinence") 
cchs20141<-subset(cchs20141, bowel_disease!="Other bowel disease") 



```


### CLEANING: SES VARIABLES 

```{r data cleaning: SES Variables}

#immigration and ethnicity
cchs20141 <- cchs20141 %>% 
  mutate(immigrant = case_when(
      SDCFIMM == 1 ~ "Immigrant",
      SDCFIMM == 2 ~ "Non-immigrant", TRUE ~ NA_character_
    ),
    cultural_racial_origin = case_when(
      SDCGCGT == 1 ~ "White",
      SDCGCGT == 2 ~ "Visible minority", TRUE ~ NA_character_
    ))

#Home ownership variable
cchs20141 <- cchs20141 %>% 
  mutate(
    home_ownership = case_when(
      DHH_OWN == 1 ~ "Own home",
      DHH_OWN == 2 ~ "Rent home"
    ))

#Marital Status Variable 
cchs20141 <- cchs20141 %>% 
  mutate(marital_status = case_when(
      DHHGMS == 1 |  DHHGMS == 2  ~ "Married or Common-law",
      DHHGMS == 3 ~ "Widowed, separated, or divorced",
      DHHGMS == 4 ~ "Single or never marrierd",
    ))

#Education 
cchs20141 <- cchs20141 %>% 
  mutate(education = case_when(
    EDUDR04 == 1 ~ "Some secondary school",
    EDUDR04 == 2 ~ "Secondary school",
    EDUDR04 == 3 ~ "Some post-secondary",
    EDUDR04 == 4 ~ "Post-secondary certifiate"
  ))

#Income
cchs20141 <- cchs20141 %>% 
  mutate(
    income_household = case_when(
      INCGHH == 1 ~ "No income or less than $20,000",
      INCGHH == 2 ~ "$20,000 to $39,999",
      INCGHH == 3 ~ "$40,000 to $59,999",
      INCGHH == 4 ~ "$60,000 to $79,999",
      INCGHH == 5 ~ "$80,000 or more"
    ))

#Employment 
  #GEN_08 "have you worked at a job or business at anytime in the past 12 months (respondents age 15-75 years)" 
cchs20141 <- cchs20141 %>% 
  mutate(employed_last_12_months = #The question is have you worked at a job or business at anytime in the past 12 months (respondents age 15-75 years)
           case_when(
             GEN_08 == 1 ~ "Yes",
             GEN_08 == 2 ~ "No or Not applicable", 
             GEN_08 == 6 ~ "No or Not applicable")  #This is not applicable 
         )#Note re:coding -- if not known, refusal, or not stated, the variable is left blank 

  #Confirming that all participants with age >=75 years are re-coded as no (vs. missing)
cchs20141$employed_last_12_months[cchs20141$DHHGAGE > 14] <- "No or Not applicable"



```

### CLEANING: CHRONIC CONDITION VARIABLES 

> Data cleaning: other chronic conditions variables  

```{r data cleaning: other chronic conditions}
 cchs20141 <- cchs20141 %>%
   mutate(
     asthma = case_when(
       CCC_031 == 1 ~ "Yes",
       CCC_031 == 2 ~ "No"
     ),
     fibromyalgia = case_when(
       CCC_041 == 1 ~ "Yes",
       CCC_041 == 2 ~ "No"
     ),
     arthritis = case_when(
       CCC_051 == 1 ~ "Yes",
       CCC_051 == 2 ~ "No"
     ),
     back_problems = case_when(
       CCC_061 == 1 ~ "Yes",
       CCC_061 == 2 ~ "No"
     ),
     high_bp = case_when(# Note: this is current blood pressure
         # There is a separate question about past blood pressure but did not include this one
       CCC_071 == 1 ~ "Yes",
       CCC_071 == 2 ~ "No"
     ),
     migraine = case_when(
       CCC_081 == 1 ~ "Yes",
       CCC_081 == 2 ~ "No"
     ),
     copd = case_when(
       CCC_091 == 1 ~ "Yes",
       CCC_091 == 2 ~ "No"
     ),
     diabetes = case_when(
       CCC_101 == 1 ~ "Yes",
       CCC_101 == 2 ~ "No"
     ),
     heart_disease = case_when(
       CCC_121 == 1 ~ "Yes",
       CCC_121 == 2 ~ "No"
     ),
     cancer = case_when(
         # Note: this is current cancer
         # There is a separate question about past blood pressure but did not include this one
       CCC_131 == 1 ~ "Yes",
       CCC_131 == 2 ~ "No"
     ),
     stomach_ulcer = case_when(
       CCC_141 == 1 ~ "Yes",
       CCC_141 == 2 ~ "No"
     ),
     stroke = case_when(
       CCC_151 == 1 ~ "Yes",
       CCC_151 == 2 ~ "No"
     ),

     urinary_incontinence = case_when(
       CCC_161 == 1 ~ "Yes",
       CCC_161 == 2 ~ "No"
     ),

     ibs = case_when(
       bowel_disease == "Irritable bowel syndrome" ~ "Yes",
       !is.na(bowel_disease) & bowel_disease != "Irritable bowel syndrome" ~ "No"
     ),

     bowel_incontinence = case_when(
       bowel_disease == "Bowel incontinence" ~ "Yes",
       !is.na(bowel_disease) & bowel_disease != "Bowel incontinence" ~ "No"
     ),

     other_bowel = case_when(
       bowel_disease == "Other bowel disease" ~ "Yes",
       !is.na(bowel_disease) & bowel_disease != "Irritable bowel syndrome" ~ "No"
     ),

     non_ibd_bowel = case_when(
       ibd == 0 & bowel_disease == "No bowel disorder" ~ "No",
       ibd == 1 ~ "No",
       ibd == 0 & !is.na(bowel_disease) ~ "Yes"
     ),

     scoliosis = case_when(
       CCC_173 == 1 ~ "Yes",
       CCC_173 == 2 ~ "No"
     ),

     chronic_fatigue = case_when(
       CCC_251 == 1 ~ "Yes",
       CCC_251 == 2 ~ "No"
     ),

     chemical_sensitivities = case_when(
       CCC_261 == 1 ~ "Yes",
       CCC_261 == 2 ~ "No"
     ),

     mood_disorder = case_when(
       CCC_280 == 1 ~ "Yes",
       CCC_280 == 2 ~ "No"
     ),

     anxiety = case_when(
       CCC_290 == 1 ~ "Yes",
       CCC_290 == 2 ~ "No"
     ),

     mood_anxiety = case_when(
       mood_disorder == "Yes" | anxiety == "Yes" ~ "Yes",
       mood_disorder == "No" & anxiety == "No" ~ "No"
     )) %>%

   rowwise() %>%
   mutate(number_chronic_conditions =
            sum(c_across(c(asthma,
                           fibromyalgia,
                           arthritis,
                           back_problems,
                           scoliosis,
                           high_bp,
                           migraine,
                           copd,
                           diabetes,
                           stomach_ulcer,
                           heart_disease,
                           cancer,
                           stroke,
                           urinary_incontinence,
                           mood_disorder,
                           anxiety
            ))=="Yes", na.rm = TRUE),
          #Creating a variable that counts the number of chronic conditions 
          number_chronic_conditions_cat = case_when(
            number_chronic_conditions == 0 ~ "0",
            number_chronic_conditions == 1 ~ "1",
            number_chronic_conditions >= 2 ~ "2 or more"
          ))
```



### CLEANING: UNMET HEALTHCARE NEED VARIABLES 

> Unmet healthcare need within the past 12 mo + WHY

```{r data cleaning: Unmet need in the past 12 mo1}
cchs20141 <- cchs20141 %>% 
  mutate(unmet_12mo = case_when(
    UCN_010 == 1 ~ "Unmet care need in the past 12 mo",
    UCN_010 == 2 ~ "No unmet care need in the past 12 mo",
  )) %>% 
    mutate(unmet_12mo_binary = case_when(
    UCN_010 == 1 ~ 1,
    UCN_010 == 2 ~ 0 ,
  )) 

cchs20141 %>% count(unmet_12mo) 
```


### CLEANING: REGULAR/IMMEDIATE HEALTHCARE USAGE 

```{r, cleaning has a regular medical doctor}

cchs20141 <- cchs20141 %>% 
  mutate(regular_doctor = case_when(
    HCU_1AA == 1 ~ 1, 
    HCU_1AA == 2 ~ 0
  )) %>%
  mutate(regular_doctor_Yes = case_when(
    HCU_1AA == 1 ~ "Yes", 
    HCU_1AA == 2 ~ "No"
  )) 

cchs20141 %>% count(regular_doctor_Yes) 

```


### CLEANING: CONSULTED HEALTHCARE PROFESSIONAL

```{r, type of healthcare professionals that one consults}

cchs20141 <- cchs20141 %>% 
  mutate(consult_family = case_when(
    CHP_03 == 1 ~ "Yes",
    CHP_03 == 2 ~ "No" 
  )) %>%
  mutate(consult_eye = case_when(
    CHP_06 == 1 ~ "Yes",
    CHP_06 == 2 ~ "No" 
  )) %>%
  mutate(consult_specialist = case_when(
    CHP_08 == 1 ~ "Yes",
    CHP_08 == 2 ~ "No" 
  )) %>%
  mutate(consult_dentist = case_when(
    CHP_14 == 1 ~ "Yes",
    CHP_14 == 2 ~ "No"
  )) %>%
  mutate(consult_chiropractor = case_when(
    CHP_16 == 1 ~ "Yes",
    CHP_16 == 2 ~ "No"
  )) %>%  mutate(consult_physiotherapist = case_when(
    CHP_18 == 1 ~ "Yes",
    CHP_18 == 2 ~ "No"
  )) %>%
  mutate(consult_psychologist = case_when(
    CHP_20 == 1 ~ "Yes",
    CHP_20 == 2 ~ "No"
  )) 


```



### TABLE 1: CONSULTED HEALTHCARE PROFESSIONAL

```{r, cross-tab by sex}

# Create the survey design object
survey_design <- svydesign(~1, data = cchs20141, weights = ~WTS_M)

# Define the variables to include in the table
vars <- c("sex", "age_group_20s_OG", "bowel_disease", "immigrant","cultural_racial_origin", "home_ownership", "marital_status", "education", 
          "number_chronic_conditions_cat", "income_household", "employed_last_12_months", "unmet_12mo",
          "regular_doctor_Yes","consult_family","consult_specialist",  "consult_psychologist", "consult_eye", "consult_dentist", "consult_chiropractor", 
          "consult_physiotherapist")

# Create the TableOne object
table_one <- svyCreateTableOne(
  vars = vars,
  strata = "ibd",
  data = survey_design,
  test = FALSE)

# Print the table
print(table_one, smd = TRUE)

# Convert the TableOne object to a data frame
table_one_df <- as.data.frame(print(table_one, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table_one_df, "table1_WEIGHTED.csv", row.names = TRUE)


####### NON-WEIGHTED 

# Create the TableOne object
table1_NoWeight <- CreateTableOne(
                        vars = vars,
                        strata = "ibd",
                        data = cchs20141,
                        test = FALSE)

# Convert the TableOne object to a data frame
table1_NoWeight_df <- as.data.frame(print(table1_NoWeight, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table1_NoWeight_df, "table1_NoWeight.csv", row.names = TRUE)


```



### TABLE 2: CONSULTED HEALTHCARE PROFESSIONAL

```{r, cross-tab by sex}

# Create the survey design object
survey_design <- svydesign(~1, data = cchs20141, weights = ~WTS_M)

# Define the variables to include in the table
vars <- c( "sex", "age_group_20s_OG", "immigrant","cultural_racial_origin", "home_ownership", "marital_status", "education", "number_chronic_conditions",
           "number_chronic_conditions_cat", "income_household", "employed_last_12_months", "regular_doctor_Yes","consult_family","consult_specialist", 
           "consult_psychologist","consult_eye", "consult_dentist", "consult_chiropractor", "consult_physiotherapist")

# Create the TableOne object
table_two <- svyCreateTableOne(
  vars = vars,
  strata = c("unmet_12mo", "ibd"),
  data = survey_design, #Using the weighted version
  test = FALSE
)

# Print the table
print(table_two, smd = TRUE)

# Convert the TableOne object to a data frame
table_two_df <- as.data.frame(print(table_two, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table_two_df, "table2_WEIGHTED.csv", row.names = TRUE)


####NON-WEIGHTED 

# Create the TableOne objects
table2_noW <- CreateTableOne(
  vars = vars,
  strata = c("unmet_12mo", "ibd"),
  data = cchs20141, #Using the non-weighted version
  test = FALSE
)

# Convert the TableOne object to a data frame
table2_noW_df <- as.data.frame(print(table2_noW, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table2_noW_df, "table2_NoWeights.csv", row.names = TRUE)

```




### CRUDE ASSOCIATION BETWEEN IBD AND UNMET NEEDS 

```{r, assessing the crude associaiton between IBD and unmet needs}

# Create the survey design object
survey_design <- svydesign(~1, data = cchs20141, weights = ~WTS_M)

#Crude logistic regression model of the association between 
model_crude <- glm(unmet_12mo_binary ~ ibd, data=cchs20141, family=binomial)

#Print the summary table of the crude model 
summary(model_crude)

#Create the confidence intervals of crude model 
exp(cbind (odd_ratio = coef (model_crude), confint(model_crude) ) )

```


### ASSESSMENT OF MULTICOLLINEARITY 

```{r, assessing the crude associaiton between IBD and unmet needs}

#Subsetting the dataframe with relevant columns to assess for multicollinearity 
myvars <- c("ibd","unmet_12mo_binary","sex", "age_group_20s", "immigrant","cultural_racial_origin", "home_ownership", "marital_status", "education", 
            "number_chronic_conditions_cat", "income_household", "employed_last_12_months", "regular_doctor_Yes","consult_specialist", 
           "consult_psychologist")
cchs20141_temp <- cchs20141[myvars]

# Fit a linear model
model <- lm(unmet_12mo_binary ~ ., data = cchs20141_temp)  

# Calculate VIF
# We defined 2.5 as the threshold for multicollinearity
vif_values <- vif(model)
print(vif_values)

```


### MODEL 1: ADJUSTED ASSOCIATION BETWEEN IBD AND UNMET NEEDS 
In this model we account for ...

```{r, creating model 1}

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_1 <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions, data=cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_1)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_1), confint(model_1) ) )


#USING A MIXED MODEL 
cchs20141$immigrant <- factor(cchs20141$immigrant, levels = c("Non-immigrant", "Immigrant"))

#Mixed_model_1 <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions_cat + marital_status + education 
#                         + home_ownership + income_household + employed_last_12_months + (1 | GEODPMF), 
#                       data = cchs20141, family = binomial, control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

Mixed_model_1 <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s_OG + immigrant + cultural_racial_origin + number_chronic_conditions_cat + 
                         income_household  + employed_last_12_months + education + marital_status + home_ownership + (1 | GEODPMF), 
                       data = cchs20141, family = binomial, control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print the summary of the model
summary(Mixed_model_1)

# table of OR (derived from esponentiaed estimates) with 95% CI
se <- sqrt(diag(vcov(Mixed_model_1))) #Calculated the standard error
tab_1 <- cbind(Est = fixef(Mixed_model_1), LL = fixef(Mixed_model_1) - 1.96 * se, UL = fixef(Mixed_model_1) + 1.96 *se)
tab_1 <- exp(tab_1) #exponentiating the table
print(round(tab_1, 2))

#Assessing the impact of respondent's health region
# Compute ICC
icc_model_1 <- icc(Mixed_model_1)

```


### MODEL 2: ADJUSTED ASSOCIATION BETWEEN IBD AND UNMET NEEDS 
In this model we account for ...

```{r, creating model 2 }

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_2 <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions + 
#                 regular_doctor_Yes + consult_specialist + consult_psychologist,
#               data=cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_2)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_2), confint(model_2) ) )

#USING A MIXED MODEL 

Mixed_model_2 <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s_OG + immigrant + cultural_racial_origin + number_chronic_conditions_cat + 
                         marital_status + education + employed_last_12_months + 
                         home_ownership + income_household  + regular_doctor_Yes + consult_specialist + consult_psychologist + (1 | GEODPMF), 
                       data = cchs20141, family = binomial, control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print the summary of the model
summary(Mixed_model_2)

# table of OR (derived from esponentiaed estimates) with 95% CI
se <- sqrt(diag(vcov(Mixed_model_2))) #Calculated the standard error
(tab_2 <- cbind(OR = fixef(Mixed_model_2), LL = fixef(Mixed_model_2) - 1.96 * se, UL = fixef(Mixed_model_2) + 1.96 *se))
tab_2 <- exp(tab_2) #exponentiating the table
print(round(tab_2, 2))

#Assessing the impact of respondent's health region
# Compute ICC
icc_model_2 <- icc(Mixed_model_2)

```


### SENSITIVITY ANALYSIS 
In the sensitivity analysis, we are evaluating the impact of self-reported mental health comorbidities.

```{r, re-defining number of chronic conditinons}

cchs20141 <- cchs20141 %>%
   rowwise() %>%
   mutate(number_chronic_conditions_SA =
            sum(c_across(c(asthma,
                           fibromyalgia,
                           arthritis,
                           back_problems,
                           scoliosis,
                           high_bp,
                           migraine,
                           copd,
                           diabetes,
                           stomach_ulcer,
                           heart_disease,
                           cancer,
                           stroke,
                           urinary_incontinence
            ))=="Yes", na.rm = TRUE), 
          
       number_chronic_conditions_cat_SA = case_when(
                  number_chronic_conditions == 0 ~ "0",
                  number_chronic_conditions == 1 ~ "1",
                  number_chronic_conditions >= 2 ~ "2 or more"
          ))


#Subsetting the dataframe with relevant columns to assess for multicollinearity 
myvars <- c("ibd","unmet_12mo_binary","sex","age_group_20s","immigrant","cultural_racial_origin","number_chronic_conditions_SA","mood_anxiety",
            "consult_family","consult_specialist","consult_psychologist")
cchs20141_temp_SA <- cchs20141[myvars]

# Fit a linear model
model <- lm(unmet_12mo_binary ~ ., data = cchs20141_temp_SA)  # 'y' is the dependent variable

# Calculate VIF
# We defined 2.5 as the threshold for multicollinearity
vif_values <- vif(model)
print(vif_values)

```

###SENSITIVITY ANALYSIS: MODEL 1

```{r, SENSITIVITY ANALYSIS: model 1}

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_1_SA <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions_SA + mood_anxiety,
#               data=cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_1_SA)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_1_SA), confint(model_1_SA) ) )


#USING A MIXED MODEL 

Mixed_model_1_SA <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s_OG + immigrant + cultural_racial_origin + number_chronic_conditions_cat_SA + marital_status + 
                            education + home_ownership + income_household + employed_last_12_months + mood_anxiety + 
                (1 | GEODPMF), data = cchs20141, family = binomial, control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print the summary of the model
summary(Mixed_model_1_SA)

# table of OR (derived from esponentiaed estimates) with 95% CI
se <- sqrt(diag(vcov(Mixed_model_1_SA))) #Calculated the standard error
(tab_1_SA <- cbind(OR = fixef(Mixed_model_1_SA), LL = fixef(Mixed_model_1_SA) - 1.96 * se, UL = fixef(Mixed_model_1_SA) + 1.96 *se))
tab_1_SA <- exp(tab_1_SA) #exponentiating the table
print(round(tab_1_SA, 2))

#Assessing the impact of respondent's health region
# Compute ICC
icc_model_1_SA <- icc(Mixed_model_1_SA)

```

###SENSITIVITY ANALYSIS: MODEL 2 

```{r, SENSITIVITY ANALYSIS: model 2}

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_2_SA <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions_SA + mood_anxiety + 
#                 regular_doctor_Yes + consult_specialist + consult_psychologist,
#               data= cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_2_SA)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_2_SA), confint(model_2_SA) ) )

#USING A MIXED MODEL 

Mixed_model_2_SA <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s_OG + immigrant + cultural_racial_origin + number_chronic_conditions_cat_SA + marital_status + 
                            education + home_ownership + income_household + employed_last_12_months + mood_anxiety + regular_doctor_Yes + 
                            consult_specialist + consult_psychologist + (1 | GEODPMF), data = cchs20141, family = binomial, 
                          control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print the summary of the model
summary(Mixed_model_2_SA)

# table of OR (derived from esponentiaed estimates) with 95% CI
se <- sqrt(diag(vcov(Mixed_model_2_SA))) #Calculated the standard error
(tab_2_SA <- cbind(OR = fixef(Mixed_model_2_SA), LL = fixef(Mixed_model_2_SA) - 1.96 * se, UL = fixef(Mixed_model_2_SA) + 1.96 *se))
tab_2_SA <- exp(tab_2_SA) #exponentiating the table
print(round(tab_2_SA, 2))

#Assessing the impact of respondent's health region
# Compute ICC
icc_model_2_SA <- icc(Mixed_model_2_SA)



```






