---
title: "Unmet Needs in the IBD Population"
author: "Gemma Postill"
date: "2022-07-29"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(tableone)
library(survey)
library(gtsummary)
library(pscl)
library(scales)
# library(MASS) # Note this package unmasks select from dplyr and wreaks havoc on the code
library(questionr)
library(pivottabler)
library(car)
library(mctest)
library(lme4)

```

# Data Preparation

> Loading the data 

```{r setup1, include=FALSE}
#This line would need to be changed 
#cchs2014 <- read_csv("C:/Users/james/OneDrive - University of Toronto/Desktop/Dr. Benchimol Work/CCHS Unmet Needs/cchs_data.csv")
cchs2014 <- read_csv("/Users/gemmapostill/Desktop/PhD/Crems22_CCHS/Data_and_Code_2014/cchs_2014_data_v4.csv")



```


### CLEANING: DESCRIPTIVE VARIABLES 


> Data cleaning: age and sex variables

```{r data cleaning: age and sex}
cchs20141 <- cchs2014 %>% 
  mutate(sex = factor(DHH_SEX,
                      levels = c(1, 2),
                      labels = c("Male", "Female")))

cchs20141 <- cchs20141 %>% 
  #This variable is grouping age in categories of 10-years 
  mutate(age_group = case_when(
    DHHGAGE == 1 | DHHGAGE == 2  ~ "< 18", #1 = 12-14, 2 = 15-17, 
    DHHGAGE == 3 | DHHGAGE == 4 | DHHGAGE == 5 ~ "18 to 29", 
    DHHGAGE == 6 | DHHGAGE == 7 ~ "30 to 39",
    DHHGAGE == 8 | DHHGAGE == 9 ~ "40 to 49",
    DHHGAGE == 10 | DHHGAGE == 11 ~ "50 to 59",
    DHHGAGE == 12 | DHHGAGE == 12 ~ "60 to 69",
    DHHGAGE == 14 | DHHGAGE == 13 ~ "70 to 79",
    DHHGAGE == 16 ~ "80 or older"
  )) 

cchs20141 <- cchs20141 %>% 
  #This variable is grouping age in categories of 20-years 
  mutate(age_group_20s = case_when(
    DHHGAGE == 1 | DHHGAGE == 2  ~ "< 18", #1 = 12-14, 2 = 15-17
    DHHGAGE == 3 | DHHGAGE == 4 | DHHGAGE == 5 | DHHGAGE == 6 | DHHGAGE == 7 ~ "18 to 39", #Note that 3 = 18-19
    DHHGAGE == 8 | DHHGAGE == 9 | DHHGAGE == 10 | DHHGAGE == 11 ~ "40 to 59",
    DHHGAGE == 12 | DHHGAGE == 12 | DHHGAGE == 14 | DHHGAGE == 13 ~ "60 to 79",
    DHHGAGE == 16 ~ "80 or older"
  )) 

cchs20141 %>% count(sex)
cchs20141 %>% count(age_group)

#Then we need to EXCLUDE the category of those with age_group = "< 18" & age_group_20s = "< 18" 
cchs20141<-subset(cchs20141, age_group!="< 18" & age_group_20s!="< 18")


```

> Data cleaning: IBD variables

```{r data cleaning: IBD1}
cchs20141 <- cchs20141 %>% 
  mutate(bowel_disease = case_when(
    CCC_171 == 1 & CCC_17A == 1 ~ "Crohn's disease",
    CCC_171 == 1 & CCC_17A == 2 ~ "Ulcerative colitis",
    CCC_171 == 1 & CCC_17A == 3 ~ "Irritable bowel syndrome",
    CCC_171 == 1 & CCC_17A == 4 ~ "Bowel incontinence",
    CCC_171 == 1 & CCC_17A == 5 ~ "Other bowel disease",
    CCC_171 == 2 ~ "No bowel disorder"
  )) %>% 
  mutate(ibd_yes = case_when(
    bowel_disease == "Crohn's disease" | bowel_disease == "Ulcerative colitis" ~ 1,
    TRUE ~ 0
  )) %>% 
  mutate(ibd = case_when(
    ibd_yes == 1 ~ "IBD Population", 
    ibd_yes == 0 ~ "General Population", 
  )) %>% 
  mutate(ibd_type = case_when(
    bowel_disease == "Crohn's disease" | bowel_disease == "Ulcerative colitis" ~ bowel_disease,
  ))

cchs20141 %>% count(bowel_disease)
cchs20141 %>% count(ibd)
cchs20141 %>% count(ibd_type)


#Then we need to EXCLUDE the category of those with:
  #bowel_disease != "Crohn's disease"
  #bowel_disease != "Ulcerative colitis"

cchs20141<-subset(cchs20141, bowel_disease!="Irritable bowel syndrome" & bowel_disease!="Bowel incontinence"
                  & bowel_disease!="Other bowel disease" )


```

> Province of Residence

```{r data cleaning: Province of Residence}
cchs20141 <- cchs20141 %>% 
  mutate(province = case_when(
    GEOGPRV == 10 ~ "Newfoundland and Labrador",
    GEOGPRV == 11 ~ "Prince Edward Island",
    GEOGPRV == 12 ~ "Nova Scotia",
    GEOGPRV == 13 ~ "New Brunswick",
    GEOGPRV == 24 ~ "Quebec",
    GEOGPRV == 35 ~ "Ontario",
    GEOGPRV == 46 ~ "Manitoba",
    GEOGPRV == 47 ~ "Saskatchewan",
    GEOGPRV == 48 ~ "Alberta",
    GEOGPRV == 59 ~ "British Columbia",
    GEOGPRV == 60 ~ "Yukon Territory",
    GEOGPRV == 61 ~ "Northwest Territories",
    GEOGPRV == 62 ~ "Nunavut"
  )) %>% 
  mutate(province_groups = case_when(
    GEOGPRV == 10 | GEOGPRV == 11 | GEOGPRV == 12 |GEOGPRV == 13 ~ "Atlantic Provinces",
    GEOGPRV == 24 ~ "Quebec",
    GEOGPRV == 35 ~ "Ontario",
    GEOGPRV == 46 | GEOGPRV == 47 | GEOGPRV == 48 ~ "Prarie Provinces",
    GEOGPRV == 59 ~ "British Columbia",
    GEOGPRV == 60 | GEOGPRV == 61 |GEOGPRV == 62  ~ "Territories"
  ))

cchs20141 %>% count(province)
cchs20141 %>% count(province_groups)
```

> Data cleaning: SES variables  
Note that SDCGCB13 (i.e., canadian_born with 1="Canadian-born" and 2="Born outside of Canada") is not in data base
Note that SDCFIMM (i.e., canadian_born with 1="Immigrant" and 2="Non-immigrant") is not in data base


```{r data cleaning: immigration and ethnicity}
cchs20141 <- cchs20141 %>% 
  mutate(immigrant = case_when(
      SDCFIMM == 1 ~ "Immigrant",
      SDCFIMM == 2 ~ "Non-immigrant", TRUE ~ NA_character_
    ),
    cultural_racial_origin = case_when(
      SDCGCGT == 1 ~ "White",
      SDCGCGT == 2 ~ "Visible minority", TRUE ~ NA_character_
    ))

```

```{r data cleaning: Home ownership variable}
cchs20141 <- cchs20141 %>% 
  mutate(
    home_ownership = case_when(
      DHH_OWN == 1 ~ "Own home",
      DHH_OWN == 2 ~ "Rent home"
    ))
```

```{r data cleaning: Marital status variable}
cchs20141 <- cchs20141 %>% 
  mutate(marital_status = case_when(
      DHHGMS == 1 |  DHHGMS == 2  ~ "Married or Common-law",
      DHHGMS == 3 ~ "Widowed, separated, or divorced",
      DHHGMS == 4 ~ "Single or never marrierd",
    ))
```

```{r data cleaning: education}
cchs20141 <- cchs20141 %>% 
  mutate(education = case_when(
    EDUDR04 == 1 ~ "Some secondary school",
    EDUDR04 == 2 ~ "Secondary school",
    EDUDR04 == 3 ~ "Some post-secondary",
    EDUDR04 == 4 ~ "Post-secondary certifiate"
  ))
```

```{r data cleaning: salary}
cchs20141 <- cchs20141 %>% 
  mutate(
    income_household = case_when(
      INCGHH == 1 ~ "No income or less than $20,000",
      INCGHH == 2 ~ "$20,000 to $39,999",
      INCGHH == 3 ~ "$40,000 to $59,999",
      INCGHH == 4 ~ "$60,000 to $79,999",
      INCGHH == 5 ~ "$80,000 or more"
    ))
```


```{r data cleaning: employment}
cchs20141 <- cchs20141 %>% 
  mutate(employed_last_12_months = 
           case_when(
             GEN_08 == 1 ~ "Yes",
             GEN_08 == 2 ~ "No"))

cchs20141 %>% count(employed_last_12_months)
```

> Data cleaning: other chronic conditions variables  



```{r data cleaning: other chronic conditions}
 cchs20141 <- cchs20141 %>%
   mutate(

     asthma = case_when(
       CCC_031 == 1 ~ "Yes",
       CCC_031 == 2 ~ "No"
     ),

     fibromyalgia = case_when(
       CCC_041 == 1 ~ "Yes",
       CCC_041 == 2 ~ "No"
     ),

     arthritis = case_when(
       CCC_051 == 1 ~ "Yes",
       CCC_051 == 2 ~ "No"
     ),

     back_problems = case_when(
       CCC_061 == 1 ~ "Yes",
       CCC_061 == 2 ~ "No"
     ),

     high_bp = case_when(
         # Note: this is current blood pressure
         # There is a separate question about past blood pressure but did not include this one
       CCC_071 == 1 ~ "Yes",
       CCC_071 == 2 ~ "No"
     ),

     migraine = case_when(
       CCC_081 == 1 ~ "Yes",
       CCC_081 == 2 ~ "No"
     ),

     copd = case_when(
       CCC_091 == 1 ~ "Yes",
       CCC_091 == 2 ~ "No"
     ),

     diabetes = case_when(
       CCC_101 == 1 ~ "Yes",
       CCC_101 == 2 ~ "No"
     ),

     heart_disease = case_when(
       CCC_121 == 1 ~ "Yes",
       CCC_121 == 2 ~ "No"
     ),

     cancer = case_when(
         # Note: this is current cancer
         # There is a separate question about past blood pressure but did not include this one
       CCC_131 == 1 ~ "Yes",
       CCC_131 == 2 ~ "No"
     ),

     stomach_ulcer = case_when(
       CCC_141 == 1 ~ "Yes",
       CCC_141 == 2 ~ "No"
     ),

     stroke = case_when(
       CCC_151 == 1 ~ "Yes",
       CCC_151 == 2 ~ "No"
     ),

     urinary_incontinence = case_when(
       CCC_161 == 1 ~ "Yes",
       CCC_161 == 2 ~ "No"
     ),

     ibs = case_when(
       bowel_disease == "Irritable bowel syndrome" ~ "Yes",
       !is.na(bowel_disease) & bowel_disease != "Irritable bowel syndrome" ~ "No"
     ),

     bowel_incontinence = case_when(
       bowel_disease == "Bowel incontinence" ~ "Yes",
       !is.na(bowel_disease) & bowel_disease != "Bowel incontinence" ~ "No"
     ),

     other_bowel = case_when(
       bowel_disease == "Other bowel disease" ~ "Yes",
       !is.na(bowel_disease) & bowel_disease != "Irritable bowel syndrome" ~ "No"
     ),

     non_ibd_bowel = case_when(
       ibd == 0 & bowel_disease == "No bowel disorder" ~ "No",
       ibd == 1 ~ "No",
       ibd == 0 & !is.na(bowel_disease) ~ "Yes"
     ),

     scoliosis = case_when(
       CCC_173 == 1 ~ "Yes",
       CCC_173 == 2 ~ "No"
     ),

     chronic_fatigue = case_when(
       CCC_251 == 1 ~ "Yes",
       CCC_251 == 2 ~ "No"
     ),

     chemical_sensitivities = case_when(
       CCC_261 == 1 ~ "Yes",
       CCC_261 == 2 ~ "No"
     ),

     mood_disorder = case_when(
       CCC_280 == 1 ~ "Yes",
       CCC_280 == 2 ~ "No"
     ),

     anxiety = case_when(
       CCC_290 == 1 ~ "Yes",
       CCC_290 == 2 ~ "No"
     ),

     mood_anxiety = case_when(
       mood_disorder == "Yes" | anxiety == "Yes" ~ "Yes",
       mood_disorder == "No" & anxiety == "No" ~ "No"
     )) %>%

   rowwise() %>%
   mutate(number_chronic_conditions =
            sum(c_across(c(asthma,
                           fibromyalgia,
                           arthritis,
                           back_problems,
                           scoliosis,
                           high_bp,
                           migraine,
                           copd,
                           diabetes,
                           stomach_ulcer,
                           heart_disease,
                           cancer,
                           stroke,
                           urinary_incontinence,
                           mood_disorder,
                           anxiety
            ))=="Yes", na.rm = TRUE),

          number_chronic_conditions_cat = case_when(
            number_chronic_conditions == 0 ~ "0",
            number_chronic_conditions == 1 ~ "1",
            number_chronic_conditions == 2 ~ "2",
            number_chronic_conditions >= 3 ~ "3 or more"
          ))
```




### CLEANING: UNMET HEALTHCARE NEED VARIABLES 

> Unmet healthcare need within the past 12 mo + WHY

```{r data cleaning: Unmet need in the past 12 mo1}
cchs20141 <- cchs20141 %>% 
  mutate(unmet_12mo = case_when(
    UCN_010 == 1 ~ "Unmet care need in the past 12 mo",
    UCN_010 == 2 ~ "No unmet care need in the past 12 mo",
  )) %>% 
    mutate(unmet_12mo_binary = case_when(
    UCN_010 == 1 ~ 1,
    UCN_010 == 2 ~ 0 ,
  )) %>% 
  mutate(unmet_area = case_when(
    UCN_010 == 1 & UCN_020A == 1 ~ 1,
    UCN_010 == 1 & UCN_020A == 2 ~ 0
  )) %>%
  mutate(unmet_available = case_when(
    UCN_010 == 1& UCN_020B == 1 ~ 1,
    UCN_010 == 1 & UCN_020B == 2 ~ 0
  )) %>%
  mutate(unmet_waiting = case_when(
    UCN_010 == 1 & UCN_020C == 1 ~ 1,
    UCN_010 == 1 & UCN_020C == 2 ~ 0
  )) %>%
  mutate(unmet_inadequate = case_when(
    UCN_010 == 1 & UCN_020D == 1 ~ 1,
    UCN_010 == 1 & UCN_020D == 2 ~ 0
  )) %>%
  mutate(unmet_cost = case_when(
    UCN_010 == 1 & UCN_020E == 1 ~ 1,
    UCN_010 == 1 & UCN_020E == 2 ~ 0
  )) %>%
  mutate(unmet_busy = case_when(
    UCN_010 == 1 & UCN_020F == 1 ~ 1,
    UCN_010 == 1 & UCN_020F == 2 ~ 0
  )) %>%
  mutate(unmet_didnt_bother = case_when(
    UCN_010 == 1 & UCN_020G == 1 ~ 1,
    UCN_010 == 1 & UCN_020G == 2 ~ 0
  )) %>%
  mutate(unmet_decided_no = case_when(
    UCN_010 == 1 & UCN_020H == 1 ~ 1,
    UCN_010 == 1 & UCN_020H == 2 ~ 0
  )) %>%
  mutate(unmet_doctor = case_when(
    UCN_010 == 1 & UCN_020I == 1 ~ 1,
    UCN_010 == 1 & UCN_020I == 2 ~ 0
  )) %>%
  mutate(unmet_other = case_when(
    UCN_010 == 1 & UCN_020J == 1 ~ 1,
    UCN_010 == 1 & UCN_020J == 2 ~ 0
  ))

cchs20141 %>% count(unmet_12mo) 
```

> Type of unmet healthcare need 

```{r data cleaning: type of unmet healthcare need}
cchs20141 <- cchs20141 %>% 
  mutate(unmet_physical = case_when(
    UCN_010 == 1 & UCN_030A == 1 ~ 1,
    UCN_010 == 1 & UCN_030A == 2 ~ 0
  )) %>%
  mutate(unmet_emotional = case_when(
    UCN_010 == 1 & UCN_030B == 1 ~ 1,
    UCN_010 == 1 & UCN_030B == 2 ~ 0
  )) %>%
  mutate(unmet_checkup = case_when(
    UCN_010 == 1 & UCN_030C == 1 ~ 1,
    UCN_010 == 1 & UCN_030C == 2 ~ 0
  )) %>%
  mutate(unmet_injury = case_when(
    UCN_010 == 1 & UCN_030D == 1 ~ 1,
    UCN_010 == 1 & UCN_030D == 2 ~ 0
  )) %>%
  mutate(unmet_type_other = case_when(
    UCN_010 == 1 & UCN_030E == 1 ~ 1,
    UCN_010 == 1 & UCN_030E == 2 ~ 0
  ))

```

> Locations where one tried to access care when unmet healthcare need 

```{r data cleaning: Unmet need in the past 12 mo}
cchs20141 <- cchs20141 %>% 
  mutate(unmet_doctor = case_when(
    UCN_010 == 1 & UCN_040A == 1 ~ 1,
    UCN_010 == 1 & UCN_040A == 2 ~ 0
  )) %>%
  mutate(unmet_clsc = case_when(
    UCN_010 == 1 & UCN_040B == 1 ~ 1,
    UCN_010 == 1 & UCN_040B == 2 ~ 0 
  )) %>%
  mutate(unmet_walkin = case_when(
    UCN_010 == 1 & UCN_040C == 1 ~ 1,
    UCN_010 == 1 & UCN_040C == 2 ~ 0 
  )) %>%
  mutate(unmet_appointment = case_when(
    UCN_010 == 1 & UCN_040D == 1 ~ 1,
    UCN_010 == 1 & UCN_040D == 2 ~ 0 
  )) %>%
  mutate(unmet_hopital_emeg = case_when(
    UCN_010 == 1 & UCN_040E == 1 ~ 1,
    UCN_010 == 1 & UCN_040E == 2 ~ 0
  )) %>%
  mutate(unmet_hospital_out = case_when(
    UCN_010 == 1 & UCN_040F == 1 ~ 1,
    UCN_010 == 1 & UCN_040F == 2 ~ 0
  )) %>%
  mutate(unmet_location_other = case_when(
    UCN_010 == 1 & UCN_040G == 1 ~ 1,
    UCN_010 == 1 & UCN_040G == 2 ~ 0
  )) 

```


### TABLES: UNMET HEALTHCARE NEED VARIABLES 

> Table of the reasons for unmet care needs in past 12 months by gender

```{r, table of reasons for / type of/ location where care saught for unmet healthcare need1}

unmet_dataset <- filter(cchs20141, unmet_12mo=="Unmet care need in the past 12 mo")

unmet_why <- unmet_dataset %>% select(ibd, unmet_area, unmet_available, unmet_waiting, 
                                            unmet_inadequate, unmet_cost, unmet_busy, unmet_didnt_bother, 
                                            unmet_decided_no, unmet_doctor, unmet_other,WTS_M) 


unmet_type <- unmet_dataset %>% select(ibd, unmet_physical, unmet_emotional, unmet_checkup, 
                                       unmet_injury, unmet_type_other,WTS_M) 


unmet_location <- unmet_dataset %>% select(ibd, unmet_doctor, unmet_clsc, unmet_walkin, unmet_appointment,
                                            unmet_hopital_emeg, unmet_hospital_out, unmet_location_other,WTS_M)


unmet_why_table <- survey::svydesign(~1, data = as.data.frame(unmet_why), weights = ~WTS_M) %>%  tbl_svysummary(by = ibd,
    statistic = list(all_continuous() ~ "{median} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_continuous() ~ 2,
                  all_categorical() ~ c(0,1)),
    label = list(unmet_area ~ "Not available in the area", 
                 unmet_available ~ "Not available at the time required",  
                 unmet_waiting ~ "Waiting time too long", 
                 unmet_inadequate ~ "Felt it would be inadequate",
                 unmet_cost ~ "Cost", unmet_busy ~ "Too busy", 
                 unmet_didnt_bother ~ "Didn't get around to it / didn't bother", 
                 unmet_decided_no ~ "Decided not to seek care", 
                 unmet_doctor ~ "Doctor didn't think it was necessary", 
                 unmet_other ~ "Other reason"),
  missing="ifany") %>%
  modify_caption("**Table 6. Description of Perceived Unmet Need within last 12 month for General Population vs. IBD Population**")

unmet_type_table <- survey::svydesign(~1, data = as.data.frame(unmet_type), weights = ~WTS_M) %>% tbl_svysummary(by = ibd,
    statistic = list(all_continuous() ~ "{median} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_continuous() ~ 2,
                  all_categorical() ~ c(0,1)),
    label = list(unmet_physical ~ "Physical health problem", 
                 unmet_emotional ~ "Emotional or mental health problem", 
                 unmet_checkup ~ "A regular check-up (including pre-natal care)",
                 unmet_injury ~ "Care of an injury", 
                 unmet_type_other ~ "Other type of care needed"),
  missing="ifany")

unmet_location_table <- survey::svydesign(~1, data = as.data.frame(unmet_location), weights = ~WTS_M) %>% tbl_svysummary(by = ibd,
    statistic = list(all_continuous() ~ "{median} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_continuous() ~ 2,
                  all_categorical() ~ c(0,1)),
    label = list(unmet_doctor ~ "Doctor's office", 
                 unmet_clsc ~ "Community health centre / CLSC", 
                 unmet_walkin ~ "Walk-in clinic", 
                 unmet_appointment ~ "Appointment clinic", 
                 unmet_hopital_emeg ~ "Hospital - emergency room", 
                 unmet_hospital_out ~ "Hospital - outpatient clinic", 
                 unmet_location_other ~ "Other location"),
  missing="ifany")  


#Stacking together interpretation tables
unmet_descriptive_table <- tbl_stack(list(unmet_why_table, unmet_type_table, unmet_location_table), group_header = c("Reason for Unmet Healthcare Need", "Type of Unmet Healthcare Need", "Location of Unmet Healthcare Need")) %>% 
    as_gt() %>% 
    gt::tab_style(style = gt::cell_text(weight = "bold"), 
    locations = gt::cells_row_groups(groups = everything()))
unmet_descriptive_table
```


### CLEANING: REGULAR/IMMEDIATE HEALTHCARE USAGE 

```{r, cleaning has a regular medical doctor}

cchs20141 <- cchs20141 %>% 
  mutate(regular_doctor = case_when(
    HCU_1AA == 1 ~ 1, 
    HCU_1AA == 2 ~ 0
  )) %>%
  mutate(regular_doctor_Yes = case_when(
    HCU_1AA == 2 & HCU_1BE == 1 ~ "Yes",
    HCU_1AA == 2 & HCU_1BE == 2 ~ "No"
  ))
cchs20141 %>% count(regular_doctor) 

```

```{r, cleaning why does not have a regular medical doctor}

cchs20141 <- cchs20141 %>% 
  mutate(non_available = case_when(
    HCU_1AA == 2 & HCU_1BA == 1 ~ 1,
    HCU_1AA == 2 & HCU_1BA == 2 ~ 0
  )) %>%
  mutate(not_new_patients = case_when(
    HCU_1AA == 2 & HCU_1BB == 1 ~ 1,
    HCU_1AA == 2 & HCU_1BB == 2 ~ 0
  )) %>%
  mutate(not_tried = case_when(
    HCU_1AA == 2 & HCU_1BC == 1 ~ 1,
    HCU_1AA == 2 & HCU_1BC == 2 ~ 0
  )) %>%
  mutate(retired = case_when(
    HCU_1AA == 2 & HCU_1BD == 1 ~ 1,
    HCU_1AA == 2 & HCU_1BD == 2 ~ 0
  )) %>%
  mutate(other_reason_health_prof = case_when(
    HCU_1AA == 2 & HCU_1BE == 1 ~ 1,
    HCU_1AA == 2 & HCU_1BE == 2 ~ 0
  ))
```

```{r, cleaning: if no regular medical doctor, regular place of care + kind of regular}

cchs20141 <- cchs20141 %>% 
  mutate(regular_care = case_when(
    HCU_1A1 == 1 ~ 1, 
    HCU_1A1 == 2 ~ 0,
  )) %>%
  mutate(kind_regular_care = case_when(
    HCU_1A2 == 1 ~ 'doctor’s office',
    HCU_1A2 == 2 ~ 'community health care centre',
    HCU_1A2 == 3 ~ 'walk-in clinic',
    HCU_1A2 == 4 ~ 'appointment clinic',
    HCU_1A2 == 5 ~ 'telephone health line',
    HCU_1A2 == 6 ~ 'hospital emergency room',
    HCU_1A2 == 7 ~ 'hospital outpatient clinic',
    HCU_1A2 == 8 ~ 'other type of immediate care' ))
    
cchs20141 %>% count(regular_care) 
```


### CLEANING: CONSULTED HEALTHCARE PROFESSIONAL

```{r, type of healthcare professionals that one consults}

cchs20141 <- cchs20141 %>% 
  mutate(consult_family = case_when(
    CHP_03 == 1 ~ "Yes",
    CHP_03 == 2 ~ "No" 
  )) %>%
  mutate(consult_eye = case_when(
    CHP_06 == 1 ~ "Yes",
    CHP_06 == 2 ~ "No" 
  )) %>%
  mutate(consult_specialist = case_when(
    CHP_08 == 1 ~ "Yes",
    CHP_08 == 2 ~ "No" 
  )) %>%
  mutate(consult_dentist = case_when(
    CHP_14 == 1 ~ "Yes",
    CHP_14 == 2 ~ "No"
  )) %>%
  mutate(consult_chiropractor = case_when(
    CHP_16 == 1 ~ "Yes",
    CHP_16 == 2 ~ "No"
  )) %>%  mutate(consult_physiotherapist = case_when(
    CHP_18 == 1 ~ "Yes",
    CHP_18 == 2 ~ "No"
  )) %>%
  mutate(consult_psychologist = case_when(
    CHP_20 == 1 ~ "Yes",
    CHP_20 == 2 ~ "No"
  )) 


```


```{r, number & location of consulatations with healthcare professionals}

#Number of consults

cchs20141 <- cchs20141 %>% 
  mutate(num_family = dplyr::case_when(
    CHPG04 <=5  ~ "1-5 visits",
    CHPG04 > 5 & CHPG04 <=10 ~ "6-10 visits",
    CHPG04 > 10 | CHPG04 == 12~ "11+ visits",
  )) %>%
  mutate(num_specialist = case_when(
    CHPG09 <=5  ~ "1-5 visits",
    CHPG09 > 5 & CHPG09 <=10 ~ "6-10 visits",
    CHPG09 == 11 | CHPG09 == 12 ~ "11+ visits"
  )) 

#Location of consults

cchs20141 <- cchs20141 %>% 
  mutate(location_family = case_when(
    CHPG05 == 1 | CHPG05 == 3 ~ 'doctor’s office or hospital outpatient clinic',
    CHPG05 == 2 ~ 'hospital emergency room',
    CHPG05 == 4 ~ 'walk-in clinic',
    CHPG05 == 5 ~ 'community health care centre or CLSC',
    CHPG05 == 6 |CHPG05 == 7 ~ 'at work, school, or other location', #includes at home
    CHPG05 == 8 ~ 'telephone consultation only',
  )) %>%
  mutate(location_specialist = case_when(
    CHPG10 == 1 | CHPG10 == 3 ~ 'doctor’s office or hospital outpatient clinic',
    CHPG10 == 2 ~ 'hospital emergency room',
    CHPG10 == 4 ~ 'walk-in clinic',
    CHPG10 == 5 ~ 'community health care centre or CLSC',
    CHPG10 == 6 | CHPG10 == 7 ~ 'at work, school, or other location', #includes at home
    CHPG10 == 8 ~ 'telephone consultation only'))




```



###DESCRIPTIVE TABLE: CONSULTED HEALTHCARE PROFESSIONAL

```{r, table of type of healthcare professionals that one consults}
professionals_consults <- cchs20141 %>% select(ibd, consult_family, consult_specialist, consult_eye, 
                                               consult_dentist, consult_chiropractor, 
                                               consult_physiotherapist, 
                                               consult_psychologist,WTS_M)

num_location_consults <- cchs20141 %>% select(ibd, consult_family, num_family, location_family, 
                                              consult_specialist, num_specialist, location_specialist,WTS_M)


table_professional_consults <- survey::svydesign(~1, data = as.data.frame(professionals_consults), weights = ~WTS_M) %>%  tbl_svysummary(by = ibd,
    statistic = list(all_continuous() ~ "{median} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_continuous() ~ 2,
                  all_categorical() ~ c(0,1)),
    label = list(consult_family ~ "Contact with a family doctor, or general practitioner", 
                 consult_specialist ~ "Contact with any other medical doctor or specialist", 
                 consult_eye ~ "Contact with an ophthalmologist or optometrist", 
                 consult_dentist ~ "Contact with a dentist, dental hygienist or orthodontist", 
                 consult_chiropractor ~ "Contact with a chiropractor", 
                 consult_physiotherapist ~ "Contact with a physiotherapist", 
                 consult_psychologist ~ "Contact with a psychologist"),
  missing="ifany")  %>%
  modify_caption("**Table Y. Contact with healthcare professional for General Population vs. IBD Population**")

table_num_location_consults <- survey::svydesign(~1, data = as.data.frame(num_location_consults), weights = ~WTS_M) %>%  tbl_svysummary(by = ibd,
    statistic = list(all_continuous() ~ "{median} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_continuous() ~ 2,
                  all_categorical() ~ c(0,1)),
    label = list(consult_family ~ "Contact with a family doctor, or general practitioner", 
                 num_family ~ "Number of consultations with family doctor", 
                 location_family ~ "Location of most recent consultations with family doctor", 
                 consult_specialist ~ "Contact with other medical doctor or specialist", 
                 num_specialist ~ "Number of consultations other specialist", 
                 location_specialist ~ "Location of most recent consultations with other specialist"), 
  missing="ifany") %>%
  bold_labels()  %>%
  modify_caption("**Table Y. Contact with healthcare professional for General Population vs. IBD Population**")
table_professional_consults



```




### TABLE 1: CONSULTED HEALTHCARE PROFESSIONAL

```{r, cross-tab by sex}

# Create the survey design object
survey_design <- svydesign(~1, data = cchs20141, weights = ~WTS_M)

# Define the variables to include in the table
vars <- c("sex", "age_group_20s", "bowel_disease", "immigrant","cultural_racial_origin", "home_ownership", "marital_status", "education", "number_chronic_conditions", 
          "number_chronic_conditions_cat", "income_household", 
          "employed_last_12_months", "regular_doctor_Yes","consult_family", "consult_eye", "consult_specialist", "consult_dentist", "consult_chiropractor", 
          "consult_physiotherapist", "consult_psychologist", "unmet_12mo")

# Create the TableOne object
table_one <- svyCreateTableOne(
  vars = vars,
  strata = "ibd",
  data = survey_design,
  test = FALSE)

# Print the table
print(table_one, smd = TRUE)

# Convert the TableOne object to a data frame
table_one_df <- as.data.frame(print(table_one, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table_one_df, "table1_WEIGHTED.csv", row.names = TRUE)


####### NON-WEIGHTED 

# Create the TableOne object
table1_NoWeight <- CreateTableOne(
                        vars = vars,
                        strata = "ibd",
                        data = cchs20141,
                        test = FALSE)

# Convert the TableOne object to a data frame
table1_NoWeight_df <- as.data.frame(print(table1_NoWeight, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table1_NoWeight_df, "table1_NoWeight.csv", row.names = TRUE)


```



### TABLE 2: CONSULTED HEALTHCARE PROFESSIONAL

```{r, cross-tab by sex}

# Create the survey design object
survey_design <- svydesign(~1, data = cchs20141, weights = ~WTS_M)

# Define the variables to include in the table
vars <- c( "age_group_20s", "sex", "cultural_racial_origin", "immigrant", "employed_last_12_months", "number_chronic_conditions_cat", "number_chronic_conditions",
           "regular_doctor_Yes","consult_family","consult_specialist", "consult_psychologist",
           "consult_eye", "consult_dentist", "consult_chiropractor", "consult_physiotherapist")

# Create the TableOne object
table_two <- svyCreateTableOne(
  vars = vars,
  strata = c("unmet_12mo", "ibd"),
  data = survey_design, #Using the weighted version
  test = FALSE
)

# Print the table
print(table_two, smd = TRUE)

# Convert the TableOne object to a data frame
table_two_df <- as.data.frame(print(table_two, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table_two_df, "table2_WEIGHTED.csv", row.names = TRUE)


####NON-WEIGHTED 

# Create the TableOne objects
table2_noW <- CreateTableOne(
  vars = vars,
  strata = c("unmet_12mo", "ibd"),
  data = cchs20141, #Using the non-weighted version
  test = FALSE
)

# Convert the TableOne object to a data frame
table2_noW_df <- as.data.frame(print(table2_noW, smd=TRUE, showAllLevels=TRUE, quote=FALSE, noSpaces=TRUE), row.names=TRUE)

# Export the data frame to a CSV file
write.csv(table2_noW_df, "table2_NoWeights.csv", row.names = TRUE)

```




### CRUDE ASSOCIATION BETWEEN IBD AND UNMET NEEDS 

```{r, assessing the crude associaiton between IBD and unmet needs}

# Create the survey design object
survey_design <- svydesign(~1, data = cchs20141, weights = ~WTS_M)

#Crude logistic regression model of the association between 
model_crude <- glm(unmet_12mo_binary ~ ibd, data=cchs20141, family=binomial)

#Print the summary table of the crude model 
summary(model_crude)

#Create the confidence intervals of crude model 
exp(cbind (odd_ratio = coef (model_crude), confint(model_crude) ) )

```


### ASSESSMENT OF MULTICOLLINEARITY 

```{r, assessing the crude associaiton between IBD and unmet needs}

#Subsetting the dataframe with relevant columns to assess for multicollinearity 
myvars <- c("ibd","unmet_12mo_binary","sex","age_group_20s","immigrant","cultural_racial_origin","number_chronic_conditions",
            "consult_family","consult_specialist","consult_psychologist")
cchs20141_temp <- cchs20141[myvars]

# Fit a linear model
model <- lm(unmet_12mo_binary ~ ., data = cchs20141_temp)  # 'y' is the dependent variable

# Calculate VIF
# We defined 2.5 as the threshold for multicollinearity
vif_values <- vif(model)
print(vif_values)

```


### MODEL 1: ADJUSTED ASSOCIATION BETWEEN IBD AND UNMET NEEDS 
In this model we account for ...

```{r, creating model 1}

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_1 <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions, data=cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_1)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_1), confint(model_1) ) )


#USING A MIXED MODEL 
cchs20141$immigrant <- factor(cchs20141$immigrant, levels = c("Non-immigrant", "Immigrant"))

Mixed_model_1 <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions + (1 | GEODPMF), 
                       data = cchs20141, family = binomial, control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print model output
print(Mixed_model_1, corr = FALSE)
se <- sqrt(diag(vcov(Mixed_model_1)))

# table of estimates with 95% CI
(tab_1 <- cbind(Est = fixef(Mixed_model_1), LL = fixef(Mixed_model_1) - 1.96 * se, UL = fixef(Mixed_model_1) + 1.96 *se))
exp(tab_1)

#Print the summary of the model
summary(Mixed_model_1)


#Assessing the impact of respondent's health region

# Extract variance components
var_components <- as.data.frame(VarCorr(Mixed_model_1))
between_region_variance <- var_components[var_components$grp == "GEODPMF", "vcov"]

# Compute ICC
icc_model_1 <- between_region_variance / (between_region_variance + (pi^2 / 3))
icc_model_1

```


### MODEL 2: ADJUSTED ASSOCIATION BETWEEN IBD AND UNMET NEEDS 
In this model we account for ...

```{r, creating model 2 }

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_2 <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions + 
#                 regular_doctor_Yes + consult_specialist + consult_psychologist,
#               data=cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_2)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_2), confint(model_2) ) )

#USING A MIXED MODEL 

Mixed_model_2 <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions + 
                 consult_family + consult_specialist + consult_psychologist + (1 | GEODPMF), data = cchs20141, family = binomial, 
                 control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print model output
print(Mixed_model_2, corr = FALSE)
se <- sqrt(diag(vcov(Mixed_model_2)))

# table of estimates with 95% CI
(tab_2 <- cbind(Est = fixef(Mixed_model_2), LL = fixef(Mixed_model_2) - 1.96 * se, UL = fixef(Mixed_model_2) + 1.96 *se))
exp(tab_2)

#Print the summary of the model
summary(Mixed_model_2)


#Assessing the impact of respondent's health region

# Extract variance components
var_components <- as.data.frame(VarCorr(Mixed_model_2))
between_region_variance <- var_components[var_components$grp == "GEODPMF", "vcov"]

# Compute ICC
icc_model_2 <- between_region_variance / (between_region_variance + (pi^2 / 3))
icc_model_2
```


### SENSITIVITY ANALYSIS 
In the sensitivity analysis, we are evaluating the impact of self-reported mental health comorbidities.

```{r, re-defining number of chronic conditinons}

cchs20141 <- cchs20141 %>%
   rowwise() %>%
   mutate(number_chronic_conditions_SA =
            sum(c_across(c(asthma,
                           fibromyalgia,
                           arthritis,
                           back_problems,
                           scoliosis,
                           high_bp,
                           migraine,
                           copd,
                           diabetes,
                           stomach_ulcer,
                           heart_disease,
                           cancer,
                           stroke,
                           urinary_incontinence
            ))=="Yes", na.rm = TRUE))


#Subsetting the dataframe with relevant columns to assess for multicollinearity 
myvars <- c("ibd","unmet_12mo_binary","sex","age_group_20s","immigrant","cultural_racial_origin","number_chronic_conditions_SA","mood_anxiety",
            "consult_family","consult_specialist","consult_psychologist")
cchs20141_temp_SA <- cchs20141[myvars]

# Fit a linear model
model <- lm(unmet_12mo_binary ~ ., data = cchs20141_temp_SA)  # 'y' is the dependent variable

# Calculate VIF
# We defined 2.5 as the threshold for multicollinearity
vif_values <- vif(model)
print(vif_values)

```


```{r, SENSITIVITY ANALYSIS: model 1}

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_1_SA <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions_SA + mood_anxiety,
#               data=cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_1_SA)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_1_SA), confint(model_1_SA) ) )


#USING A MIXED MODEL 

Mixed_model_1_SA <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions_SA + mood_anxiety + 
                (1 | GEODPMF), data = cchs20141, family = binomial, control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print model output
print(Mixed_model_1_SA, corr = FALSE)
se <- sqrt(diag(vcov(Mixed_model_1_SA)))

# table of estimates with 95% CI
(tab_1_SA <- cbind(Est = fixef(Mixed_model_1_SA), LL = fixef(Mixed_model_1_SA) - 1.96 * se, UL = fixef(Mixed_model_1_SA) + 1.96 *se))
exp(tab_1_SA)

#Print the summary of the model
summary(Mixed_model_1_SA)


#Assessing the impact of respondent's health region

# Extract variance components
var_components <- as.data.frame(VarCorr(Mixed_model_1_SA))
between_region_variance <- var_components[var_components$grp == "GEODPMF", "vcov"]

# Compute ICC
icc_model_1_SA <- between_region_variance / (between_region_variance + (pi^2 / 3))
icc_model_1_SA

```



```{r, SENSITIVITY ANALYSIS: model 2}

#build the model (adjusting for sex, age group, immigrant status, ethnicity, employment, number of chronic conditions)
#model_2_SA <- glm(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions_SA + mood_anxiety + 
#                 regular_doctor_Yes + consult_specialist + consult_psychologist,
#               data= cchs20141, family=binomial)

#Print the summary table of the crude model 
#summary(model_2_SA)

#Create the confidence intervals of crude model 
#exp(cbind (odd_ratio = coef (model_2_SA), confint(model_2_SA) ) )

#USING A MIXED MODEL 

Mixed_model_2_SA <- glmer(unmet_12mo_binary ~ ibd + sex + age_group_20s + immigrant + cultural_racial_origin + number_chronic_conditions_SA + mood_anxiety + 
                 consult_family + consult_specialist + consult_psychologist + (1 | GEODPMF), data = cchs20141, family = binomial, 
                 control =glmerControl(optimizer = "bobyqa"), nAGQ = 10)

#Print model output
print(Mixed_model_2_SA, corr = TRUE)
se <- sqrt(diag(vcov(Mixed_model_2_SA)))

# table of estimates with 95% CI
(tab_2<- cbind(Est = fixef(Mixed_model_2_SA), LL = fixef(Mixed_model_2_SA) - 1.96 * se, UL = fixef(Mixed_model_2_SA) + 1.96 *se))
exp(tab_2)

#Print the summary of the model
summary(Mixed_model_2_SA)

#Assessing the impact of respondent's health region

# Extract variance components
var_components <- as.data.frame(VarCorr(Mixed_model_2_SA))
between_region_variance <- var_components[var_components$grp == "GEODPMF", "vcov"]

# Compute ICC
icc_model_2_SA <- between_region_variance / (between_region_variance + (pi^2 / 3))
icc_model_2_SA

```






